Pool Model - A Mathematical Framework for Modelling Bacterial Consortia in Food Science

### 2025-12-08 paper/pool_model.tex
add fernando to authors
add orcids; remove website of jonas
Add corrections from Christian

\title{Pool Model - A unified Mathematical\\ Modeling Framework for Food Science}
\author[1]{Polina Gaindrik\orcidlink{0000-0003-0928-3951}}
\author[1]{Jonas Pleyer\fnref{fn1}}
\author[1]{Christian Fleck\corref{cor1}}
\ead{christian.fleck@fdm.uni-freiburg.de}

\fntext[fn1]{jonas.pleyer.org}
\cortext[cor1]{Corresponding Author\\ \textit{E-mail address:} christian.fleck@fdm.uni-freiburg.de}
\affiliation[1]{
    organization={Freiburg Center for Data Analysis and Modeling},
    addressline={Ernst-Zermelo Str. 1},
    city={Freiburg im Breisgau},
    postcode={79104},
    country={Germany},
}

@@

Stronger and quicker changes will lead to a higher rate $\gamma$.

@@

    \dot{\gamma} = \Gamma |\dot{T}|-\delta \gamma.
We assumed that the direction of the temperature change does not matter, i.e., a change from $T=2$ to $T=12$ has the same effect as a change from $T=14$ to $T=4$.
$\Gamma$ is a scaling factor and $\delta$ is the relaxation time for the temperature disturbance.

@@

    \gamma(t) = \Gamma \int_0^t |\dot{T}(t^{\prime})|e^{-\delta (t-t^{\prime})}dt^{\prime}.
After some manipulation of this equation we find:
    \gamma(t) = \Gamma\delta\left |T[t]-T[0]e^{-\delta t}-\delta
        \int\limits_0^t T(t^{\prime})e^{-\delta (t-t^{\prime})}dt^{\prime}\right |.
This representation of the rate $\gamma(t)$ has the advantage that it only involves $T(t)$ and does not require the calculation of the derivative of $T$.

@@

    \gamma(t) &=& \Gamma\sum_{i=1}^n \left |\Delta T_i \right |\theta(t-t_i)e^{-\delta(t-t_i)}.

>>>>

\title{Pool Model - A Mathematical Framework for Modelling Bacterial Consortia in Food Science}
\author[1]{Polina Gaindrik\orcidlink{0000-0003-0928-3951}}
\author[1]{Jonas Pleyer\orcidlink{0000-0003-0928-3951}}
\author[2]{Fernando Pérez Rodríguez}
\author[1]{Christian Fleck\orcidlink{0000-0002-6371-4495}\corref{cor1}}
\ead{christian.fleck@fdm.uni-freiburg.de}

\cortext[cor1]{Corresponding Author\\ \textit{E-mail address:} christian.fleck@fdm.uni-freiburg.de}
\affiliation[1]{
    organization={Freiburg Center for Data Analysis and Modeling},
    addressline={Ernst-Zermelo Str. 1},
    city={Freiburg im Breisgau},
    postcode={79104},
    country={Germany},
}
\affiliation[2]{
    organization={Department of Food Science and Technology, UIC Zoonosis y Enfermedades Emergentes ENZOEM, ceiA3, Universidad de Córdoba},
    addressline={Campus Rabanales s/n. Darwin building- Annex},
    city={Córdoba},
    postcode={14014},
    country={Spain},
}

@@

Stronger and quicker changes will lead to a higher rate $\gamma$, but the effect will relax to zero after some time.

@@

    \dot{\gamma} = \Gamma f\left(\dot{T}\right)-\delta \gamma.
The dimensionless function $f$, with $f(0)=0$ and $f: \mathbb{R}\to\mathbb{R}_{\ge 0}$, evaluates the effect of a temperature change, i.e., a change from $T=2$ to $T=12$ could have a different effect as a change from $T=14$ to $T=4$. $\Gamma$ is a scaling factor and $\delta$ is the relaxation time for the temperature disturbance.

@@

    \gamma(t) = \Gamma \int_0^t f\left(\dot{T}(t^{\prime})\right)e^{-\delta (t-t^{\prime})}dt^{\prime}.
One possible choice for $f$ reads:
f(x)&=\begin{cases}\alpha x& x\ge 0\\ -(1-\alpha) x& x < 0,\end{cases}
$\alpha\in [0,1]$. For $\alpha=0.5$ it follows $f(x)=\frac{1}{2}|x|$. Using this choice for $f$ we find after some manipulation:
    \gamma(t) = \Gamma\alpha\left [T(t)-T(0)e^{-\delta t}-\delta
        \int\limits_0^t T(t^{\prime})e^{-\delta (t-t^{\prime})}dt^{\prime}\right ],
for $\dot{T}\ge 0, \,\forall t$ and
    \gamma(t) = \Gamma(1-\alpha)\left [T(0)e^{-\delta t}-T(t)+\delta
        \int\limits_0^t T(t^{\prime})e^{-\delta (t-t^{\prime})}dt^{\prime}\right ].
for $\dot{T}< 0, \,\forall t$. 
This representation of the rate $\gamma(t)$ has the advantage that it only involves $T(t)$ and does not require the calculation of a derivative.

@@

    \gamma(t) &=& \Gamma\sum_{i=1}^n f\left(\Delta T_i \right)\theta(t-t_i)e^{-\delta(t-t_i)}.
